Bird Audio Detection Challenge

### 2017-03-18 paper/main.tex
Added entry to table
More edits to paper

\author{Weilian Song\\ 
University of Kentucky\\ 
{\tt\small weilian.song@uky.edu}
% For a paper whose authors are all at the same institution, omit the
% following lines up until the closing ``}''.  Additional authors and
% addresses can be added with ``\and'', just like the second author.  To
% save space, use either the email address or home page, not both
\and Nathan Jacobs\\ 
University of Kentucky\\ 
{\tt\small jacobs@cs.uky.edu}
}

@@

	Acoustic detection of birds in the wild is useful in many biological
	scenarios, including estimation of bird population and species-specific
	acoustic monitoring. Most of current methods are either manual or
	semi-automatic, requiring labor-intensive processing/tuning. We hereby
	present three alternative methods, as part of the Bird Audio Detection
	Challenge, hosted by Queen Mary University of London. Two methods are our
	own submission, with the third one as the winning solution submitted by
	Thomas Grill. All methods utilize deep neural networks for automatic,
	tuning-free bird presence detection.

@@

In section 2, we discuss some of the competition details, including the
datasets used, task description and the evaluation metrics. Section 3
presents three methods for bird acoustic detection, two of which are our
submission to the challenge, and the third one being the best performing
method submitted by Thomas Grill. Section 4 discusses some experiments and
techniques used to improve the peformance. Section 5 displays the
evaluation results of the three methods, and section 6 presents future
possible work with this challenge.

@@

For the challenge, teams have to design a system that takes a short audio
clip, and returns a binary label denoting the presence/absence of bird
sound of any species. Teams are also encouraged to output a continuous,
probability score in the range of 0 to 1 for evaluation purposes.

@@

\textbf{Chernobyl:} Part of TREE (Transfer-Exposure-Effects) research
project. Consists of audio in the Chernobyl Exclusion Zone.

Each dataset contains audio clips of roughly 10 seconds. For the challenge,
all datasets' audio files are provided, but only \textbf{freefield} and
\textbf{Warblr}'s labels are given. Held-out dataset consists primarily of
the \textbf{Chernobyl} dataset, along with some from \textbf{freefield} and
\textbf{Warblr}.

@@

Our first method resembles basic characteristics of SoundNet by Aytar et
al.\cite{soundnet} and Residual Net by He et al.\cite{residual_net}. The
network is fully convolutional, with one-dimensional convolutions followed
by nonlinearities (ReLU) layers.  This network is suitable for audio
signals, as it is invariant to translation and allows stacking of
convolutional layers for detecting high-level features.

@@

Our second method utilizes a series of pooling and convolutional layers to
extract sets of features from the input audio at different resolutions, and
concatenate them together before some final convolutional layers.

Network can be divided into three stages. In the first stage, we first
create two copies of the input, one reshaped to 100,000 by 4, and the other
reshaped to 200,000 by 2. \(x_{100,000*4}\) is then passed through a single
convolutional layer while \(x_{200,000*2}\) is passed through a series of 5
pooling layers, followed by a convolutional layer at the end of each
pooling layer. Note here that the input of one pooling layer is the output
of the previous pooling layer, the outputs of the convolutional layers
serve as features extracted at input of different resolution.

In the second stage, we combined the outputs of the 6 convolutional layers
in Stage 1 into one feature vector. First, we find the maximum for each
convolutional output. We denote them as the energy for each convolutional
layer. We then concatenate them with their corresponding convolutional
output, pass each concatenated feature vector through a pooling layer of
specified size, and concatenate all 6 feature vectors into one. Due to
different lengths of output for the 6 convolutional layers in stage 1, we
pass them through pooling layers of different sizes to ensure that they are
of the same dimension for concatenation.

@@

The given labeled data from both datasets is split into 10 folds. Within
each fold, number of positive and negative samples are balanced. During
training, number of samples loaded from each datasets are also balanced, to
ensure no bias.

@@

For method 3, code base is available online at...

@@

We deploy data augmentation as our only pre-processing technique for method
1 and 2. Given one base sound clip, positive or negative (bird
present/absent), and another negative sound clip, we concatenate them along
the y-axis, with the base clip weighted by a factor \(r\) between 0.25 and
1, and the negative sound clip weighted by a factor \(1-r\). So we have:
\begin{equation}
	x_{augmented}=r*x_{base} + (1-r)*x_{negative}
\end{equation}
If \(r\) equals to 1, then there would be no augmentation, and if \(r\)
equals to 0.25, only 25\% of the base sound clip will be preserved. We
sample \(r\) on a positively skewed interval (0.25,1) to ensure most bird
sounds in augmented clips are still detectable.

@@

This post-processing technique averages prediction scores from many
sessions. Due to the random cropping of audio signals during the data
loading phase, different evaluation of the same audio clip will yield
slightly different results. Therefore we predict 100 sets of scores for the
held-out challenge dataset, and average them to obtain one set of scores
that was submitted.

For method 3, instead of averaging sets of output scores from the same
model, different models' outputs were averaged. Please refer to Grill's
code base for more details.

@@

\begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|p{2cm}| }
	\hline
	\multicolumn{2}{|c|}{AUC Scores}\\
	\hline
	Method & AUC Score\\
	\hline
	Method 1 & 77.9\\
	Method 2 & ??.?\\
	Method 3 & 88.7\\
	\hline
\end{tabular}

>>>>

\author{
	Weilian Song, Tawfiq Salem, Nathan Jacobs\\ 
	University of Kentucky\\ 
	{\tt\small weilian.song@uky.edu, tawfiq.salem@uky.edu, jacobs@cs.uky.edu}
}

@@

	Acoustic detection of birds in the wild is useful in many biological
	scenarios, including estimation of bird population and species-specific
	acoustic monitoring. Most of current methods are either manual or
	semi-automatic, requiring labor-intensive processing/tuning. We hereby
	present three alternative methods, as part of the Bird Audio Detection
	Challenge, hosted by Queen Mary University of London. Two methods are our
	own submission, with the third one as the winning solution submitted by
	Thomas Grill. All methods utilize deep neural networks for automatic,
	tuning-free bird presence detection, with the first two methods
	emphasizing network architectures, and the third one utilizing multiple
	signal preprocessing techniques.

@@

In section 2, we discuss some of the competition details, including the
datasets used, task description and the evaluation metrics. Section 3
presents three methods for bird acoustic detection, two of which are our
submission to the challenge, and the third one being the best performing
method submitted by Thomas Grill. Section 4 discusses some experiments and
techniques used to improve the peformance. Section 5 displays the
evaluation results of the three methods, and section 6 presents future
possible work from this challenge.

@@

For the challenge, teams have to design a system that takes a short audio
clip, and returns a binary label denoting the presence/absence of bird
sound of any species. Teams are also encouraged to output a floating-point,
probability score in the range of 0 to 1 for evaluation purposes.

@@

\textbf{Chernobyl:} Part of TREE (Transfer-Exposure-Effects) research
project. Consists of ambient audio in the Chernobyl Exclusion Zone.

Each dataset contains audio clips of roughly 10 seconds. For the challenge,
all datasets' audio files are provided, but only \textbf{freefield} and
\textbf{Warblr}'s labels are given. Held-out challenge dataset consists
primarily of the \textbf{Chernobyl} dataset, along with some from
\textbf{freefield} and \textbf{Warblr}.

@@

Our first method resembles basic characteristics of SoundNet by Aytar et
al.\cite{soundnet} and Residual Net by He et al.\cite{residual_net}. The
network is fully convolutional, with one-dimensional convolutions followed
by non-linearities (ReLU) layers.  This network is suitable for audio
signals, as it is invariant to translation and allows stacking of
convolutional layers for detecting high-level features.

@@

Our second method utilizes a series of pooling and convolutional layers to
extract features from the input audio at different temporal resolutions,
and concatenate them together before some final convolutional layers.

Network can be divided into three stages. In the first stage, we first
create two copies of the input, one reshaped to 100,000 by 4, and the other
reshaped to 200,000 by 2. \(x_{100,000*4}\) is then passed through a single
convolutional layer while \(x_{200,000*2}\) is passed through a series of 5
averaging pooling layers, followed by a convolutional layer at the end of
each pooling layer. Note here that the input of one pooling layer is the
output of the previous pooling layer, the outputs of the convolutional
layers serve as features extracted at input of different temporal
resolution.

In the second stage, we combined the outputs of the 6 convolutional layers
in Stage 1 into one feature vector. First, we find the maximum for each
convolutional output. We denote them as the energy for each convolutional
layer. We then concatenate them with their corresponding convolutional
output, pass each concatenated feature vector through a pooling layer of
specified size, and concatenate all 6 pooled feature vectors into one. Due
to different lengths of output from the 6 convolutional layers in stage 1,
we pass them through pooling layers of different sizes to ensure that they
are of the same dimension for concatenation.

@@

The given labeled data from both datasets is split into 10 folds. Within
each fold, number of positive(bird present) and negative(bird absent)
samples are balanced. During training, number of samples loaded from each
datasets are also balanced, to ensure no bias.

@@

For method 3, a brief description and code base is available online
at\cite{winning_blog}

@@

We deploy data augmentation as our only pre-processing technique for method
1 and 2. Given one base sound clip, positive or negative (bird
present/absent), and another negative sound clip, we concatenate them along
the y-axis, with the base clip weighted by a factor \(r\) between 0.25 and
1, and the negative sound clip weighted by a factor \(1-r\). So we have:
\begin{equation}
	x_{augmented}=r*x_{base} + (1-r)*x_{negative}
\end{equation}
If \(r\) equals to 1, then there would be no augmentation, and if \(r\)
equals to 0.25, only 25\% of the base sound clip will be preserved. We
sample \(r\) on a positively skewed interval (0.25,1) to ensure bird sounds
in positive labels are not diminished.

@@

This post-processing technique averages prediction scores from many
sessions. Due to the random cropping of audio signals during the data
loading phase, different evaluation of the same audio clip will yield
slightly different results. Therefore we predict 100 sets of scores for the
held-out challenge dataset, and average them to capture the true scores for
submission.

For method 3, instead of averaging sets of output scores from the same
model, different models' outputs were averaged. Please refer to Grill's
code base\cite{winning_blog} for more details.

@@

\begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|p{2cm}| }
	\hline
	\multicolumn{2}{|c|}{AUC Scores}\\
	\hline
	Method & AUC Score\\
	\hline
	Method 1 & 77.9\\
	Method 2 & 78.8\\
	Method 3 & 88.7\\
	\hline
\end{tabular}
